---
title: "Log-normal distribution"
author: "Ainesh Sewak"
output: html_document
---

```{r knitr_options, include=FALSE}
library(knitr)
opts_chunk$set(fig.width=12, fig.height=4, fig.path='RmdFigs/',
               warning=FALSE, message=FALSE)
if(!require("ggplot2")) install.packages("ggplot2", repos="http://cran.us.r-project.org")
```

### Preliminaries

Let's start by setting a seed so the charts are reproducible.

```{r set_seed}
set.seed(353)
```

Also we will be using `ggplot2` for any plotting. As such you need to first install the package (if you haven't already done so). Type (within R) `install.packages("ggplot2")`

You then load the R/ggplot2 package using the `require` function:

```{r load_ggplot2}
require(ggplot2)
```

### Log-normal distribution background

If a random variable $X$ is log-normally distributed then $\ln(X)$ has a normal distribution. ie.

If $X \sim \text{Lognormal}(\mu, \sigma)$ then $\ln(X) \sim  \mathcal{N}(\mu, \sigma)$  

<br>
 
The log-normal distribution has density given by

$f(x) = \frac{1}{\sqrt{2\pi}\sigma x} \text{exp}(\frac{-(\log(x) - \mu)^2}{2 \sigma^2})$

where $\mu$ and $\sigma$ are the mean and standard deviation of the logarithm.

<br>

The arithmetic mean and the arithmetic variance of a log-normally distributed variable $X$ is given by

${E(X) = \text{exp}(\mu + \frac{1}{2} \sigma^2)}{}$, 

${Var(X) = \text{exp}(2\mu + \sigma^2)(\text{exp}(\sigma^2) - 1)}$

### Create a random sample

Next we can use the `rlnorm()` function to generate samples from a log-normal distribution $X \sim \text{Lognormal}(\mu = 1, \sigma = 0.25)$

```{r generate_lnX}
mu <- 1
sigma <- 0.25
x <- rlnorm(n = 100, meanlog = mu, sdlog = sigma)
```

The sample mean $\bar{X}$ is `r mean(x)` and the sample variance $\ s^2$ is `r var(x)`.

The theoretical mean of the distribution $X$ is `r exp(1 + 0.5*sigma^2)` and the variance is `r exp(2*mu + sigma^2) * (exp(sigma^2) - 1)`.

Note: As $lim_{n\to\infty}$, $\bar{X} \to \mu$ and $\ s^2 \to \sigma^2$.


### Histogram on a linear scale

We can use the `ggplot` function to generate a histogram for $X$.

```{r generate_hist, fig.height=8}
ggplot(data = data.frame(x), aes(x)) + 
  geom_histogram(binwidth = 0.25, fill = I("grey"), col = I("black")) +
  scale_y_continuous(breaks = seq(0,15,2)) + 
  labs(title = "Histogram of sample from a log-normal distribution", x = "X", y = "Observation Count")
```

### Histogram on a log-scale

It is also interesting to take a look at our distribution on a log scale. This is because it should essentially look like a normal distribution on this scale.

```{r generate_hist_logscale, fig.height=8}

ggplot(data = data.frame(x), aes(x)) + 
  geom_histogram(binwidth = 0.1, fill = I("grey"), col = I("black")) +
  scale_x_continuous(trans='log') + 
  labs(title = "Histogram of sample from a log-normal distribution (log scale)", x = "X (log scale)", y = "Observation Count")
```
